---
title: "Assignment_7b_SunilS"
author: "Sunil S"
date: "20 September 2017"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls(all=TRUE))

setwd("E:/INSOFE_CPEE/Wk6/09092017_Day1/201700909_Batch32_BLR_CSE7302c_LogReg_NaiveBayesLAB/20170715_Batch29_CSE7302c_NB_Activity.nb")
```

##Read file
```{r echo =TRUE, include=FALSE}
flight <- read.csv("FlightDelays.csv")

```

## libraries
```{r echo =TRUE, include=FALSE}
library(e1071)
```

## check data
```{r echo =TRUE, include=TRUE}
head(flight)
str(flight)
summary(flight)

```
## pre processing data, converting int to factor and binning DEP_TIME
```{r echo =TRUE, include=FALSE}

flight$Weather<-as.factor(flight$Weather)
flight$DAY_WEEK<-as.factor(flight$DAY_WEEK)
flight$Flight.Status<-as.factor(flight$Flight.Status)

flight$levels<-ifelse(flight$DEP_TIME>=600 & flight$DEP_TIME<=1200,"level1",
               ifelse(flight$DEP_TIME>=1200 & flight$DEP_TIME<=1800,"level2",
                      ifelse(flight$DEP_TIME>=1800 & flight$DEP_TIME<=2100,"level3","level4")))

flight$DEP_TIME <- NULL
flight$levels <- as.factor(flight$levels)
```

## checking distribution of Levels in flight status
```{r echo =TRUE}
plot(flight$Flight.Status,
     col=c("Green","Red"),
     main="Distribution of Levels in Flight Status",
     ylab="No. of levels of 0 & 1",
     xlab="Different levels in Flight Status"
     )
```
## splitting data as train and test
```{r echo = TRUE}
set.seed(123)
rows=seq(1,nrow(flight),1)
trainRows=sample(rows,(70*nrow(flight))/100)
train = flight[trainRows,]
test = flight[-trainRows,]
dim(train)
dim(test)
```


```{r echo=TRUE}
table(flight$Flight.Status)
str(flight)
```


## checking distribution of levels
```{r echo=TRUE, include=TRUE}
cat("\n Percentage of Not Delayed=",length(which(flight$Flight.Status==0))/nrow(flight))
cat("\n Percentage of Delayed=",length(which(flight$Flight.Status==1))/nrow(flight))
```
## Target distribution in train data and test data
```{r echo=TRUE, include=TRUE}
cat("\n Percentage of Not Delayed in train data=",length(which(train$Flight.Status==0))/nrow(train))
cat("\n Percentage of Delayed in train data=",length(which(train$Flight.Status==1))/nrow(train))
cat("\n Percentage of Not Delayed in test data=",length(which(test$Flight.Status==0))/nrow(test))
cat("\n Percentage of Delayed in test data=",length(which(test$Flight.Status==1))/nrow(test))
```

#Building model using Naive Bayes
```{r echo=TRUE, include=TRUE}
library(e1071)
model_nb = naiveBayes(Flight.Status ~., data = train)
model_nb$tables

```
## predict the flight status on test data
```{r echo=TRUE, include=TRUE}
nb_pred = predict(model_nb, newdata = test)


```
## confusion matrix
```{r echo =True, include = TRUE}
Confusionmatrix = table(test$Flight.Status,nb_pred,dnn=c("actual","predicted"))
Confusionmatrix
TP=Confusionmatrix[2,2];  
TN=Confusionmatrix[1,1];  
FP=Confusionmatrix[1,2];  
FN=Confusionmatrix[2,1];  
cat('\n TP=',TP)
cat('\n TN=',TN)
cat('\n FP=',FP)
cat('\n FN=',FN)
```

##checking accuracy, Precision, Sensitivity, Specificity
```{r echo =TRUE, include=TRUE}
accuracy<-(TP+TN)/(TP+TN+FP+FN)
cat('\n Accuracy=',accuracy)
Precision<-(TP)/(TP+FP)
cat('\n Precision=', Precision)
Sensitivity<-(TP)/(TP+FN)
cat('\n Sensitivity=', Sensitivity)
Specificity=(TN)/(TN+FP)
cat('\n Specificity=',Specificity)
    
```

##inspecting Flight data
```{r echo =TRUE, include=TRUE}
str(flight)
flight$Flight.Status = as.numeric(flight$Flight.Status)

flight$CARRIER = as.numeric(flight$CARRIER)
sd_Carrier = sd(flight$CARRIER)
sd_Carrier
cor(x = flight$CARRIER,y = flight$Flight.Status) #-0.102295

flight$DEST = as.numeric(flight$DEST)
sd_Dest = sd(flight$DEST)
sd_Dest
cor(x = flight$DEST,y = flight$Flight.Status) #-0.094

flight$ORIGIN = as.numeric(flight$ORIGIN)
sd_Origin = sd(flight$ORIGIN)
sd_Origin
cor(x = flight$ORIGIN,y = flight$Flight.Status) #0.05666

flight$Weather = as.numeric(flight$Weather)
sd_Weather = sd(flight$Weather)
sd_Weather
cor(x = flight$Weather,y = flight$Flight.Status) #0.2472

flight$DAY_WEEK = as.numeric(flight$DAY_WEEK)
sd_DayWeek = sd(flight$DAY_WEEK)
sd_DayWeek
cor(x = flight$DAY_WEEK,y = flight$Flight.Status) #-0.0407

flight$levels = as.numeric(flight$levels)
sd_levels = sd(flight$levels)
sd_levels
cor(x = flight$levels,y = flight$Flight.Status) #0.1208

#as correlation with Flight.Status is least with DEST try model by removing DEST.

```
```{r echo = TRUE, include = TRUE}
train1 = train
train1$DEST = NULL
model_nb1 = naiveBayes(Flight.Status ~., data = train1)
model_nb$tables

test1 = test
test1$DEST = NULL
nb_pred1 = predict(model_nb1, newdata = test1)

Confusionmatrix = table(test1$Flight.Status,nb_pred1,dnn=c("actual","predicted"))
Confusionmatrix
TP=Confusionmatrix[2,2];  
TN=Confusionmatrix[1,1];  
FP=Confusionmatrix[1,2];  
FN=Confusionmatrix[2,1];  
cat('\n TP=',TP)
cat('\n TN=',TN)
cat('\n FP=',FP)
cat('\n FN=',FN)

accuracy<-(TP+TN)/(TP+TN+FP+FN)
cat('\n Accuracy=',accuracy)
Precision<-(TP)/(TP+FP)
cat('\n Precision=', Precision)
Sensitivity<-(TP)/(TP+FN)
cat('\n Sensitivity=', Sensitivity)
Specificity=(TN)/(TN+FP)
cat('\n Specificity=',Specificity)
```
## We observe that after dropping DEST, there is increase in Accuracy and Specificity of the model, while Precision and Sensitivity has decreased. For this model, it is required to be precisie in predicting Delay. So Sensitivity and Precision is most important.Removing DEST does not help.

 
## Trying PCA
```{r echo =TRUE, include = TRUE}
train2 = train;
str(train2)
train2$CARRIER =as.numeric(train2$CARRIER)
train2$DEST =as.numeric(train2$DEST)
train2$ORIGIN =as.numeric(train2$ORIGIN)
train2$Weather =as.numeric(train2$Weather)
train2$DAY_WEEK =as.numeric(train2$DAY_WEEK)
train2$levels =as.numeric(train2$levels)

train2.predictors <- train2[,-6]
str(train2.predictors)

scaled.Predictors <- scale(train2.predictors)
scaled.Predictors
pca <- princomp(scaled.Predictors)
summary(pca)

```
## 5 components are enough as they describe 90% of the variation. Modelling using PCA.
```{r echo =TRUE, include =TRUE}
train2$Flight.Status =as.numeric(train2$Flight.Status)
pc_traindata<-data.frame(pca$scores[,1:5],train2$Flight.Status)

mod_pc_lm<-lm(train2.Flight.Status ~.,data=pc_traindata)
summary(mod_pc_lm)

test2 = test[,-6]
test2$CARRIER =as.numeric(test2$CARRIER)
test2$DEST =as.numeric(test2$DEST)
test2$ORIGIN =as.numeric(test2$ORIGIN)
test2$Weather =as.numeric(test2$Weather)
test2$DAY_WEEK =as.numeric(test2$DAY_WEEK)
test2$levels =as.numeric(test2$levels)

str(test2)
pc_test<-predict(pca,newdata=test2[1:6])
pc_test<-data.frame(pc_test[,1:6])

#Apply new model on this transformed test data
pred_test_pc<-predict(mod_pc_lm,newdata=pc_test)

test$Flight.Status = as.numeric(test$Flight.Status)
#calculate evaluation metrics
regr.eval(test$Flight.Status,pred_test_pc)

```
## With least errrors, PCA seems to improve the model.